### 2016-08-11 paper.Rmd
updated stuff

---
author: "Joe Brew"
date: "August 11, 2016"
fig_height: 2.6
fig_width: 4
output:
  html_document:
    toc: true
    toc_float: true
    theme: yeti
---

---
title: "Impact of LAIV on School Absenteeism"
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F}
# No scientific notation
options(scipen=999)

# Packages 
library(xtable)
library(knitr)
library(dplyr)
library(maptools)
library(gpclib)
library(ggplot2)
library(RColorBrewer)
library(tidyr)
library(plotly)

# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = FALSE, 
               cache = FALSE)

# Source helper functions
source('lib/functions.R')

# Read in data
ab <- read_and_clean_absenteeism_data()
im <- read_and_clean_immunization_data()

# Join and clean up data
students <- ab %>%
  dplyr::select(id, lunch, race, school, grade, dob) %>%
  filter(!duplicated(id)) %>%
  dplyr::rename(roster_lunch = lunch,
         roster_race = race,
         roster_school = school,
         roster_grade = grade,
         roster_dob = dob)
students <-
  left_join(x = students %>%
              mutate(id = as.character(id)),
            y = im,
            by = c('id' = 'student_id')) %>%
  mutate(consent_form_return =
           ifelse(is.na(consent_form_return), 
                  FALSE,
                  consent_form_return),
         consent_form_yes = 
           ifelse(is.na(consent_form_yes),
                  FALSE,
                  consent_form_yes),
         vaccine = 
           ifelse(is.na(vaccine),
                  FALSE,
                  vaccine)) %>%
  dplyr::select(id,
                roster_lunch,
                roster_race,
                roster_school,
                roster_grade,
                roster_dob,
                consent_form_return,
                consent_form_yes,
                vaccine) %>%
  dplyr::rename(lunch = roster_lunch,
                race = roster_race,
                school = roster_school,
                grade = roster_grade,
                dob = roster_dob)
```

Some text here.